### 2025-06-11 lecturenotes_aima104/ch3_aima104.tex
eigen-values

@@

@@

\item Let $A= \begin{bmatrix}
  0 & 1 \\
  -1 & 0
\end{bmatrix}$. Find the eigenvalue-eigenvector pairs. Explore the geometric effect of letting $\displaystyle x^{(k)}=Ax^{(k-1)}$ and $k=0,1,2,...$.
\end{enumerate}

\section{Diagonalization}
\begin{mdframed}
  A square matrix A is said to be diagonalizable if, there exists an invertible matrix $P$ and a diagonal matrix $D$, such that $$A = PDP^{-1}.$$ If $A = PDP^{-1}$ then $A$ is said to be similar to $D$.
\end{mdframed}

>>>>

\subsection{Exercise}
\begin{enumerate}
\item Is $
  \begin{bmatrix}
    1 \\ 4
  \end{bmatrix}$ an eigenvector of $\begin{bmatrix} -3 & 1  \\  -3 & -8  \\ \end{bmatrix}$ ?
\item Is $\lambda =2$ a eigenvalue of
  a. $\begin{bmatrix}
    3 & 2  \\
    3 & 8  \\
  \end{bmatrix}$ ?
\end{enumerate}

\begin{mdframed}

@@

\end{mdframed}

@@

\item Find a basis for the eigenspace corresponding to each eigenvalue.
  \begin{multicols}{3}
    a. $\begin{bmatrix}
      5 & 0  \\
      2 & 1  \\
    \end{bmatrix}$

    b. $\begin{bmatrix}
      10 & -9 \\
      4 & -2  \\
    \end{bmatrix}$
    \columnbreak

    c. $\begin{bmatrix}
      4 & 0 & 0 \\
      -2 & -1 & 0 \\
      -2 & 0 & 1
    \end{bmatrix}$
  \end{multicols}

\item Let $A= \begin{bmatrix}
  0 & 1 \\
  -1 & 0
\end{bmatrix}$. Find the eigenvalue-eigenvector pairs. Explore the geometric effect of letting $\displaystyle x^{(k)}=Ax^{(k-1)}$ and $k=0,1,2,...$.\

\item Prove that $A$ and $A^{t}$ have the same eigenvalues.
\end{enumerate}

\subsection{Important Results}
\begin{theorem}
  The eigenvalues of a triangular matrix are the entries on its main diagonal.
\end{theorem}

\begin{theorem}
  If $v_1, ..., v_r$ are eigenvectors that corresponds to distinct eigenvalues $\lambda _1, ..., \lambda _r$of an $n \times n$ matrix $A$, then the set $\{v_1, ..., v_r\}$ is linearly independent.
  \end{theorem}

  \begin{theorem}
    Zero is an eigenvalue of $A$ if and only if $A$ is not invertible.
  \end{theorem}
\section{Diagonalization}
\begin{mdframed}
  A square matrix A is said to be diagonalizable if, there exists an invertible matrix $P$ and a diagonal matrix $D$, such that $$A = PDP^{-1}.$$ If $A = PDP^{-1}$ then $A$ is said to be similar to $D$.
\end{mdframed}

### 2025-06-11 lecturenotes_aima203/ch4_aima203.tex
eigen-values

\section{Divided Differences}
Given $(n+1)$ points, $(x_0,y_0), \, (x_1,y_1), \, ..., \, (x_n,y_n)$ of $x$ and $y(x)$, with $x$ values equally spaced, the divided differences of order $1,2,...,n$ are defined by the relations:
\begin{align*}
  [x_0,x_1] &= \frac{y_1 - y_0}{x_1 - x_0} \\[1mm]
  [x_0,x_1,x_2] &= \frac{[x_1, x_2]-[x_0,x_1]}{x_2 - x_0} \\[1mm]
  ... \\
  [x_0,x_1,x_2,...,x_n] &= \frac{[x_1, x_2,...,x_n]-[x_0,x_1,...,x_{n-1}]}{x_n - x_0}
\end{align*}
It is easy to see that, $\displaystyle [x_0,x_1] = \frac{y_1 - y_0}{x_1 - x_0} = \frac{y_0}{x_0 - x_1} + \frac{y_1}{x_1 - x_0} = [x_1,x_0]$.
Again,
\begin{align*}
  [x_0,x_1,x_2] &= \frac{[x_2, x_0]-[x_0,x_1]}{x_2 - x_0} \\[1mm]
                &= \frac{1}{x_2 - x_0} \left [ \frac{y_2- y_1}{x_2 - x_1} - \frac{y_1 - y_0}{x_1 - x_0} \right ] \\[1mm]
                  &= \frac{1}{x_2 - x_0} \left [ \frac{y_2}{x_2 - x_1} -y_1 \left ( \frac{1}{x_2-x_1} + \frac{1}{x_1 - x_0} \right ) + \frac{y_0}{x_1 - x_0} \right ] \\[1mm]
  &= \frac{y_0}{(x_0-x_1)(x_0-x_2)} + \frac{y_1}{(x_1-x_0)(x_1-x_2)} + \frac{y_2}{(x_2-x_0)(x_2-x_1)}
\end{align*}
\begin{figure}[h]
  \centering
\includegraphics[width=13cm, height=5cm]{divided_for.png}
\end{figure}
\subsection{Newton's General Interpolation Formula}
We have,  $\displaystyle [x,x_0] = \frac{y - y_0}{x - x_0} \implies y = y_0 + (x-x_0)[x,x_0]$ \\[1mm]
And $\displaystyle [x,x_0,x_1] =  \frac{[x, x_0]-[x_0,x_1]}{x - x_1} \implies = [x,x_0] = [x_0,x_1] + (x-x_1)[x,x_0,x_1]$ \\[1mm]
From these two relations we get,
\begin{equation}
  \label{ninter}
  y = y_0 + (x-x_0)[x_0,x_1] + (x-x_0)(x-x_1)[x,x_0,x_1]
\end{equation}
Similarly we can get $[x,x_0,x_1] =  [x_0,x_1,x_2] + (x-x_2)[x,x_0,x_1,x_2]$ and so on...
Proceeding this way we get,
\begin{equation}
  \label{eq:ngeneral}
  \begin{gathered}
    y = y_0 + (x-x_0)[x_0,x_1] + (x-x_0)(x-x_1)[x_0,x_1,x_2] + \\[1mm]
    (x-x_0)(x-x_1)(x-x_2)[x_0,x_1,x_2,x_3] + ... \\[1mm]
    + (x-x_0)(x-x_1)(x-x_2)...(x-x_n)[x,x_0,x_1,...,x_n]
  \end{gathered}
  \end{equation}
  This Equation~\ref{eq:ngeneral} is called the Newton's general interpolation formula with divided differences.

>>>>

\section{Divided Differences}
Given $(n+1)$ points, $(x_0,y_0), \, (x_1,y_1), \, ..., \, (x_n,y_n)$ of $x$ and $y(x)$, with $x$ values equally spaced, the divided differences of order $1,2,...,n$ are defined by the relations:
\begin{align*}
  [x_0,x_1] &= \frac{y_1 - y_0}{x_1 - x_0} \\[1mm]
  [x_0,x_1,x_2] &= \frac{[x_1, x_2]-[x_0,x_1]}{x_2 - x_0} \\[1mm]
  ... \\
  [x_0,x_1,x_2,...,x_n] &= \frac{[x_1, x_2,...,x_n]-[x_0,x_1,...,x_{n-1}]}{x_n - x_0}
\end{align*}
It is easy to see that, $\displaystyle [x_0,x_1] = \frac{y_1 - y_0}{x_1 - x_0} = \frac{y_0}{x_0 - x_1} + \frac{y_1}{x_1 - x_0} = [x_1,x_0]$.
Again,
\begin{align*}
  [x_0,x_1,x_2] &= \frac{[x_2, x_0]-[x_0,x_1]}{x_2 - x_0} \\[1mm]
                &= \frac{1}{x_2 - x_0} \left [ \frac{y_2- y_1}{x_2 - x_1} - \frac{y_1 - y_0}{x_1 - x_0} \right ] \\[1mm]
                  &= \frac{1}{x_2 - x_0} \left [ \frac{y_2}{x_2 - x_1} -y_1 \left ( \frac{1}{x_2-x_1} + \frac{1}{x_1 - x_0} \right ) + \frac{y_0}{x_1 - x_0} \right ] \\[1mm]
  &= \frac{y_0}{(x_0-x_1)(x_0-x_2)} + \frac{y_1}{(x_1-x_0)(x_1-x_2)} + \frac{y_2}{(x_2-x_0)(x_2-x_1)}
\end{align*}
\begin{figure}[h]
  \centering
\includegraphics[width=13cm, height=5cm]{divided_for.png}
\end{figure}
\subsection{Newton's General Interpolation Formula}
We have,  $\displaystyle [x,x_0] = \frac{y - y_0}{x - x_0} \implies y = y_0 + (x-x_0)[x,x_0]$ \\[1mm]
And $\displaystyle [x,x_0,x_1] =  \frac{[x, x_0]-[x_0,x_1]}{x - x_1} \implies [x,x_0] = [x_0,x_1] + (x-x_1)[x,x_0,x_1]$ \\[1mm]
From these two relations we get,
\begin{equation}
  \label{ninter}
  y = y_0 + (x-x_0)[x_0,x_1] + (x-x_0)(x-x_1)[x,x_0,x_1]
\end{equation}
Similarly we can get $[x,x_0,x_1] =  [x_0,x_1,x_2] + (x-x_2)[x,x_0,x_1,x_2]$ and so on...
Proceeding this way we get,
\begin{equation}
  \label{eq:ngeneral}
  \begin{gathered}
    y = y_0 + (x-x_0)[x_0,x_1] + (x-x_0)(x-x_1)[x_0,x_1,x_2] + \\[1mm]
    (x-x_0)(x-x_1)(x-x_2)[x_0,x_1,x_2,x_3] + ... \\[1mm]
    + (x-x_0)(x-x_1)(x-x_2)...(x-x_n)[x,x_0,x_1,...,x_n]
  \end{gathered}
  \end{equation}
  This Equation~\ref{eq:ngeneral} is called the Newton's general interpolation formula with divided differences.
